Variable Density Tesseroids: Gravity fields calculation in spherical coordinates using variable densities in depth

### 2017-10-17 manuscript/manuscript.tex
Edited determination of D with linear density in manuscript
Modified manuscript with new density-based discretization algorithm

In order to overcome this, we developed a complementary discretization algorithm (it works together with the previous one) that takes into account the variation of the density inside the tesseroid.
Basically, we will subdivide any tesseroid on the radius coordinate on the depth at which the \emph{maximum density variation} takes place, creating two new tesseroids which we use to compute the desired gravity field by applying the previous adaptive discretization algorithm and the GLQ approximation.

First we need to define a simple density that approximates the one of the tesseroid: a linear function that assumes the same values of the original density on the tesseroid's top and bottom:

\begin{equation}
    \rho_0(r') =
    \frac{\rho(R_2) - \rho(R_1)}{R_2 - R_1} (r' - R_1) + \rho(R_1).
\end{equation}

Then we will define the \emph{maximum density variation} as the maximum difference between this density $\rho_0(r')$ and the original density $\rho(r')$ and obtain the radius coordinate $R_\text{max}$ at which this \emph{maximum} takes place:

\begin{equation}
    \text{max}\{ | \rho(r') - \rho_0(r') | \} = 
    | \rho(R_\text{max}) - \rho_0(R_\text{max}) |
\end{equation}

Finally, we subdivide the original tesseroid at $R_\text{max}$ creating two new tesseroids with the same latitude and longitude boundaries.

This density-based discretization can be applied several times, creating a new pair of subdivisions on each iteration and therefore increasing the accuracy of the resulting gravity field.
Because each tesseroid is subdivided in two every time we apply this discretization, the number of subdivisions increases exponentially in power of two:

\begin{equation}
    \text{Number of subdivisions} =
    \sum\limits_{m=0}^M 2^m,
\end{equation}

\noindent where $M$ is the order of the discretization, i.e. how many times the density-based discretization algorithm is applied on one tesseroid.

It's worth notice that for $M=0$ or for linear densities, the density-based discretization algorithm will not be applied.
In the first case because no subdivisions will be made, and in the second one due to the fact that the absolute difference of densities is equal to zero on every depth, thus no having a meaningful maximum.

@@

\begin{enumerate}
    \item Applies the density-based discretization algorithm in case of variable density producing subdivisions of the original tesseroid,
    \item Applies the modified adaptive discretization algorithm for each subdivision, generating a set of smaller tesseroids,
    \item Computes the desired gravity field for every small tesseroid through the GLQ approximation (equation \ref{eq:glq-var-dens}).
\end{enumerate}

@@

>>>>

In order to overcome this, we developed a complementary discretization algorithm that takes into account the variation of the density inside the tesseroid.
Basically, it subdivides any tesseroid on the radius coordinate on the depth at which the \emph{maximum density variation} takes place.

Specifically, the algorithm consists in the following steps:

\begin{enumerate}
\renewcommand{\theenumi}{(\arabic{enumi})}
    \item Normalise the density function to the [0, 1] interval,
    \item Compute the absolute difference between the normalised density and a linear function that assumes the same values on the top and bottom radii that this normalised density.
    \item If the maximum of the absolute difference is greater that a predefined delta ratio ($\delta$), then the tesseroid will be subdivided on the radius at which this maximum takes place. Otherwise, the tesseroid won't be subdivided.
\end{enumerate}

First we need to normalise the density function such as it assumes the values of 0 and 1 on the boundary radii ($R_1$, $R_2$), respectively, depending at which radius the original density function is maximum.
We call this normalised density $\rho_n(r')$ and we can calculate it as follows:

\begin{equation}
    \rho_n(r') = \frac{\rho(r') - \rho_1}{\rho_2 - \rho_1},
\end{equation}
\noindent where
\begin{equation}
    \rho_1 = \text{min}\{ \rho(R_1), \rho(R_2) \}, \quad
    \rho_2 = \text{max}\{ \rho(R_1), \rho(R_2) \}.
\end{equation}

Secondly, we must define a linear function $\rho_l(r')$ that assumes the same values of the normalised density $\rho_n(r')$ on the boundary radii:

\begin{equation}
    \rho_l(r') =
    \frac{ \rho_n(R_2) - \rho_n(R_1) }{ R_2 - R_1 } (r' - R_1) + \rho_n(R_1),
\end{equation}

\noindent and compute the absolute difference between this linear function and the normalised density:

\begin{equation}
    \Delta \rho (r') = | \rho_n(r') - \rho_l(r') |
\end{equation}

Finally, if the maximum of the absolute difference if greater than a predefined delta ratio $\delta$ we will divide the tesseroid on the radius at which this maximum takes place ($R_\text{max}$):

\begin{equation}
    \text{max}\{ \Delta \rho(r') \} > \delta
    \label{eq:delta-density}
\end{equation}
\begin{equation}
    \text{max}\{ \Delta \rho(r') \} = \Delta \rho(R_\text{max}),
\end{equation}

\noindent obtaining a pair of smaller tesseroids with the same longitude and latitude boundaries.

In case that the original density function assumes the same value at both boundary radii, the procedure is the same except that the constants $\rho_1$ and $\rho_2$ are redefined as follows:

\begin{equation}
    \rho_1 = \text{min}\{ \rho(r') \}, \quad
    \rho_2 = \text{max}\{ \rho(r') \}.
\end{equation}

If on the contrary the inequation \ref{eq:delta-density} is not held, then the tesseroid will not be subdivided.
The same happens if the density function is constant or if the tesseroid radius dimension is bellow the threshold of 1mm.
On the first case due to the fact that the density-based discretization algorithm is not needed, and on the second one because smaller tesseroids will not introduce meaningful increases in the accuracy of the computation.

This algorithm can be subsequently applied to every subdividing tesseroid, until the inequation \ref{eq:delta-density} is not held or the subdivisions are smaller than 1mm.
This two conditions guarantee that the algorithm is finite and does not fall into an infinite loop. \todo{chequear esta oracion}
Once this conditions are met, every subdividing tesseroid is subjected to the modified adaptive discretization algorithm and the GLQ in order to compute the gravity field generated by it.

It's easy to see that the delta ratio controls how many times the tesseroids will be subdivided based on the variation of the density function and therefore it controls the accuracy involving the numerical error that the density variation introduces.
This raise the need of determine a maximum value of $\delta$ that ensures an acceptable accuracy.
We will perform this determination with the same methodology as the one applied to the $D$ ratio: compare the numerical model with the analytical solutions for a spherical shell.

@@

\begin{enumerate}
    \renewcommand{\theenumi}{(\arabic{enumi})}
    \item Applies the density-based discretization algorithm in case of variable density producing subdivisions of the original tesseroid,
    \item Applies the modified adaptive discretization algorithm for each subdivision, generating a set of smaller tesseroids,
    \item Computes the desired gravity field for every small tesseroid through the GLQ approximation (equation \ref{eq:glq-var-dens}).
\end{enumerate}

@@

It's worth to remember that the density-based discretization algorithm is not applied for linear density function comparisons, thus the adaptive discretization algorithm is the only mechanism that controls the accuracy of the approximation.
